Macroscopic Salt Rejection through Electrostatically Gated Porous Graphene

### 2019-01-09 paper/paper.tex
change symbol for graphene radius

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
with those of bare PCTE membrane, suggesting the transport resistance
of PCTE $R_{\mathrm{PCTE}}$, is dominant over that of PG,
$R_{\mathrm{G}}^{0}$. Using a simple circuit model, we estimate the
ratio $\delta$ between $R_{\mathrm{G}}^{0}$ and $R_{\mathrm{PG}}$,
such that $\delta = R_{\mathrm{PCTE}}/R_{\mathrm{G}}^{0}$, to be ~2,
which is in good agreement with the value derived from the
experimental data in \Fig{} \ref{fig:2}b (details see Supplementary Section SXX).
\todo[inline]{Make sure that the values used are correct.}

@@

The numerical procedure described above allows us to resolve the
concentration and electric potential profiles near a graphene
nanopore for a given $V_{\mathrm{G}}$. Since the ionic flux is driven
by both $\psi$ and $c$ fields (see Eq. \eqref{eq:pnp} and \eqref{eq:mu}),
we focus on the electrochemical potential $\mu_{i}$, which represents
the combined driving force, to reveal its dependence on the applied
$V_{\mathrm{G}}$. Following the same system considered in \Fig{}
\ref{fig:3}b, the calculated axisymmetric electric potential $\psi$
and the relative electrochemical potentials,
$\Delta \mu_{\mathrm{K^{+}}}$ and $\Delta \mu_{\mathrm{Cl^{-}}}$, for
$V_{\mathrm{G}}$ = 0.75 V are shown in \Fig{} \ref{fig:4}a and
\ref{fig:4}b, respectively. The reference point of the electrochemical
potentials is set at the bulk phase in the HCR. As expected, the
electric potential reaches the maximum at the graphene surface (with
$\psi_{\mathrm{G}}$ of $\sim$100 mV) and decays towards the bulk
solution phase, forming an EDL surrounding the graphene surface. Since
the Debye length $\lambda_{\mathrm{D}}$ is comparable to the
pore radius $R_{\mathrm{G}}$, the electric potential at the nanopore
center remains at $\sim$25 mV, comparable to the thermal energy at
room temperature ($k_{\mathrm{B}}T$ = 26 meV). Consequently, it is
evident that the potential barrier is sufficient to modulate the
diffusive flux.

We further reveal the cation and anion transport pathways by looking
into their electrochemical potentials (\Fig{} \ref{fig:4}b). By
increasing $V_{\mathrm{G}}$, a more positive $\psi_{\mathrm{G}}$ on
graphene surface reduces the cation concentration at the pore edge due
to the electrostatic interactions, thereby increasing its
concentration gradient at the pore center. On the other hand, the
anion concentration near the pore edge increases exponentially, such
that the concentration gradient at the pore center becomes
small. (also see Supplementary \Fig{} S4). We also observe that
$\Delta \mu_{+}$ is dominated by the concentration, while
$\Delta \mu_{-}$ is much less than $\Delta \mu_{+}$ due to the balance
between the diffusion and drift potentials (see Supplementary \Fig{}
S5). The observations confirm the distinct ionic transport pathways
upon applying a positive $V_{\mathrm{G}}$: pore center for cations and
pore edge for anions. \Fig{} \ref{fig:4}c presents the calculated
\textit{z}-component of the cation and anion fluxes through the pore
at the graphene plane (z = 0), $\boldsymbol{J}_{z}$ ,as a function of
the normalized radius, $r/R_{\mathrm{G}}$ , where $r$ is the radial
coordinate and $R_{\mathrm{G}}$ is the pore radius, at different
$V_{\mathrm{G}}$ values. Note that $\boldsymbol{J}_{z}$ is negative
because the LCR is placed at z$<$0 in the simulation box. At
$V_{\mathrm{G}}$ = 0 (pure diffusion), the anion and cation flux
profiles are identical, with a higher flux near the pore edge, which
is expected, as the concentration gradient is higher. By gradually
increasing $V_{\mathrm{G}}$, both fluxes are reduced throughout the
pore, while the degree of reduction is more pronounced at the pore
edge for the cation, and at the pore center for the anion, following
the scenarios we discussed above. Accordingly, the integrated fluxes
across the pore, $|\boldsymbol{J}_{z}|$, are reduced with
$V_{\mathrm{G}}$ (\Fig{} \ref{fig:4}d). A total flux reduction of up
to 80\% is predicted. Another interesting finding here is that, by
increasing $V_{\mathrm{G}}$, the nanopore transport preferentially
allows cations over anions, or in other words, the anion flux is more
reduced with $V_{\mathrm{G}}$ (see \Fig{} \ref{fig:4}c), known as the
ion selectivity of graphene nanopore \cite{Rollings_2016}. In our setup, the imbalance between cation and anion
fluxes near the graphene nanopore is compensated by the EDL formation
at the counter-electrode in LCR, which maintains the electroneutrality
of the system. Notably, since atomically thin graphene samples are
used, the mechanism of salt rejection is clearly distinct from that in
capacitive deionization \cite{Biesheuvel_2010}, where electrodes with large
surface area are required.

The above mechanistic findings, nevertheless, do not fully clarify the
experimentally observed reduction of diffusive flux upon
gating. Indeed, the same physical mechanism also governs the ionic
transport through an IFET, in which a nonzero electric potential at
the nanopore center usually results in an increase of ionic
conductivity \cite{Nam_2009,Lee_2015,Feng_2016}. To this end, we
further increased the graphene surface potential $\psi_{\mathrm{G}}$
considering the same system (Supplementary Section 6). Intriguingly,
by increasing $\psi_{\mathrm{G}}$ larger than 400 mV, the ionic flux
starts to increase, reversing the trend observed at the low
$\psi_{\mathrm{G}}$ regime, due to (i) the drift flux become dominant
over diffusion flux and (ii) the concentration near the pore edge
(prefered path) greatly increases (details see Supplementary Section
S7, Supplementary Figs. S8 and S9). This level of $\psi_{\mathrm{G}}$
cannot be reached experimentally in graphene membrane, as this
requires a $V_{\mathrm{G}}$ larger than 2.0 V, which triggers the
electrochemical reactions \cite{Toh_2011}. On the other hand, in an
IFET, the electric potential on the pore wall is considerably higher,
equivalent to $V_{\mathrm{G}}$ due to the metallic nature of gate
electrode (having an infinitely-large quantum capacitance). We
therefore conclude that the quantum capacitance-induced nonlinear
damping in the surface potential results in the observed salt
rejection.  Following the above discussions, we further investigate
the physical limits for the biased graphene nanopores. Specifically,
the salt rejection characteristics are controlled by: (i) the overlap
of EDL inside the nanopore, essentially controlled by two length
scales of $\lambda_{\mathrm{D}}$ and $R_{\mathrm{G}}$, and (ii) the
graphene surface potential $\psi_{\mathrm{G}}$ controlled by
$V_{\mathrm{G}}$ following Eq. \eqref{eq:Vg}. To this end, we
calculate $\psi_{\mathrm{G}}$ as a function of
$\lambda_{\mathrm{D}} / R_{\mathrm{G}}$ and $V_{\mathrm{G}}$ (\Fig{}
\ref{fig:5}a). Clearly, when the bulk concentration $c_{0}$ increases,
a higher $V_{\mathrm{G}}$ is required to reach the same level of
$\psi_{\mathrm{G}}$. Consequently, as revealed in \Fig{} \ref{fig:5}b,
the calculated salt rejection factor $\xi$ increases with both
$\lambda_{\mathrm{D}} / R_{\mathrm{G}}$ and $V_{\mathrm{G}}$, in line
with the experimentally observed $\xi - c_{0}$ dependence. We also
find the overall trends for $\psi_{\mathrm{G}}$ and $\xi$ with respect
to $\lambda_{\mathrm{D}}/R_{\mathrm{G}}$ and $V_{\mathrm{G}}$ are very
similar. More importantly, based on our theoretical prediction, with
$V_{\mathrm{G}}$ = 1.25 V and $\lambda_{\mathrm{D}} / R_{\mathrm{G}}$
= 1.0, a high value of $\xi$ of $>$ 0.85 can be achieved.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/fig5.pdf}
  \caption{\textbf{The Debye length effect on the degree of salt
      rejection}. \textbf{a}. Calculated contour plot of the surface
    potential of graphene $\psi_{\mathrm{G}}$ as a function of
    $\lambda_{\mathrm{D}}/R_{\mathrm{G}}$ and $V_{\mathrm{G}}$. When
    the Debye length reduces, a higher $V_{\mathrm{G}}$ is required to
    maintain the same surface $\psi_{\mathrm{G}}$
    level. \textbf{b}. Calculated 2D contour plot of theoretically
    predicted $\xi$ as a function of
    $\lambda_{\mathrm{D}}/R_{\mathrm{G}}$ and
    $V_{\mathrm{G}}$. Clearly, a higher degree of salt rejection can
    be achieved by having a higher $V_{\mathrm{G}}$ and
    $\lambda_{\mathrm{D}}/R_{\mathrm{G}}$}
  \label{fig:5}
\end{figure}

>>>>

Next, we carried out the same experiments using the PG-covered PCTE
membrane (PG-PCTE), without electrostatic gating. The measured
diffusive flux values, $\boldsymbol{J}_{\mathrm{PG0}}$ (green bars in
\Fig{} \ref{fig:2}b), remain at the same order of magnitude compared
with those of bare PCTE membrane, suggesting that the transport resistance
of PCTE $R_{\mathrm{PCTE}}$, is dominant over that of PG without gating,
$R_{\mathrm{G}}^{0}$. Using a simple circuit model, we estimate the
ratio $\delta$ between $R_{\mathrm{G}}^{0}$ and $R_{\mathrm{PCTE}}$,
such that $\delta = R_{\mathrm{PCTE}}/R_{\mathrm{G}}^{0}$, to be ~2,
which is in good agreement with the value derived from the
experimental data in \Fig{} \ref{fig:2}b (details see Supplementary Section SXX).
\todo[inline]{Make sure that the values used are correct.}

@@

The numerical procedure described above allows us to resolve the
concentration and electric potential profiles near a graphene
nanopore for a given $V_{\mathrm{G}}$. Since the ionic flux is driven
by both $\psi$ and $c$ fields (see Eq. \eqref{eq:pnp} and \eqref{eq:mu}),
we focus on the electrochemical potential $\mu_{i}$, which represents
the combined driving force, to reveal its dependence on the applied
$V_{\mathrm{G}}$. Following the same system considered in \Fig{}
\ref{fig:3}b, the calculated axisymmetric electric potential $\psi$
and the relative electrochemical potentials,
$\Delta \mu_{\mathrm{K^{+}}}$ and $\Delta \mu_{\mathrm{Cl^{-}}}$, for
$V_{\mathrm{G}}$ = 0.75 V are shown in \Fig{} \ref{fig:4}a and
\ref{fig:4}b, respectively. The reference point of the electrochemical
potentials is set at the bulk phase in the HCR. As expected, the
electric potential reaches the maximum at the graphene surface (with
$\psi_{\mathrm{G}}$ of $\sim$100 mV) and decays towards the bulk
solution phase, forming an EDL surrounding the graphene surface. Since
the Debye length $\lambda_{\mathrm{D}}$ is comparable to the
pore radius $r_{\mathrm{G}}$, the electric potential at the nanopore
center remains at $\sim$25 mV, comparable to the thermal energy at
room temperature ($k_{\mathrm{B}}T$ = 26 meV). Consequently, it is
evident that the potential barrier is sufficient to modulate the
diffusive flux.

We further reveal the cation and anion transport pathways by looking
into their electrochemical potentials (\Fig{} \ref{fig:4}b). By
increasing $V_{\mathrm{G}}$, a more positive $\psi_{\mathrm{G}}$ on
graphene surface reduces the cation concentration at the pore edge due
to the electrostatic interactions, thereby increasing its
concentration gradient at the pore center. On the other hand, the
anion concentration near the pore edge increases exponentially, such
that the concentration gradient at the pore center becomes
small. (also see Supplementary \Fig{} S4). We also observe that
$\Delta \mu_{+}$ is dominated by the concentration, while
$\Delta \mu_{-}$ is much less than $\Delta \mu_{+}$ due to the balance
between the diffusion and drift potentials (see Supplementary \Fig{}
S5). The observations confirm the distinct ionic transport pathways
upon applying a positive $V_{\mathrm{G}}$: pore center for cations and
pore edge for anions. \Fig{} \ref{fig:4}c presents the calculated
\textit{z}-component of the cation and anion fluxes through the pore
at the graphene plane (z = 0), $\boldsymbol{J}_{z}$ ,as a function of
the normalized radius, $r/r_{\mathrm{G}}$ , where $r$ is the radial
coordinate and $r_{\mathrm{G}}$ is the pore radius, at different
$V_{\mathrm{G}}$ values. Note that $\boldsymbol{J}_{z}$ is negative
because the LCR is placed at z$<$0 in the simulation box. At
$V_{\mathrm{G}}$ = 0 (pure diffusion), the anion and cation flux
profiles are identical, with a higher flux near the pore edge, which
is expected, as the concentration gradient is higher. By gradually
increasing $V_{\mathrm{G}}$, both fluxes are reduced throughout the
pore, while the degree of reduction is more pronounced at the pore
edge for the cation, and at the pore center for the anion, following
the scenarios we discussed above. Accordingly, the integrated fluxes
across the pore, $|\boldsymbol{J}_{z}|$, are reduced with
$V_{\mathrm{G}}$ (\Fig{} \ref{fig:4}d). A total flux reduction of up
to 80\% is predicted. Another interesting finding here is that, by
increasing $V_{\mathrm{G}}$, the nanopore transport preferentially
allows cations over anions, or in other words, the anion flux is more
reduced with $V_{\mathrm{G}}$ (see \Fig{} \ref{fig:4}c), known as the
ion selectivity of graphene nanopore \cite{Rollings_2016}. In our setup, the imbalance between cation and anion
fluxes near the graphene nanopore is compensated by the EDL formation
at the counter-electrode in LCR, which maintains the electroneutrality
of the system. Notably, since atomically thin graphene samples are
used, the mechanism of salt rejection is clearly distinct from that in
capacitive deionization \cite{Biesheuvel_2010}, where electrodes with large
surface area are required.

The above mechanistic findings, nevertheless, do not fully clarify the
experimentally observed reduction of diffusive flux upon
gating. Indeed, the same physical mechanism also governs the ionic
transport through an IFET, in which a nonzero electric potential at
the nanopore center usually results in an increase of ionic
conductivity \cite{Nam_2009,Lee_2015,Feng_2016}. To this end, we
further increased the graphene surface potential $\psi_{\mathrm{G}}$
considering the same system (Supplementary Section 6). Intriguingly,
by increasing $\psi_{\mathrm{G}}$ larger than 400 mV, the ionic flux
starts to increase, reversing the trend observed at the low
$\psi_{\mathrm{G}}$ regime, due to (i) the drift flux become dominant
over diffusion flux and (ii) the concentration near the pore edge
(prefered path) greatly increases (details see Supplementary Section
S7, Supplementary Figs. S8 and S9). This level of $\psi_{\mathrm{G}}$
cannot be reached experimentally in graphene membrane, as this
requires a $V_{\mathrm{G}}$ larger than 2.0 V, which triggers the
electrochemical reactions \cite{Toh_2011}. On the other hand, in an
IFET, the electric potential on the pore wall is considerably higher,
equivalent to $V_{\mathrm{G}}$ due to the metallic nature of gate
electrode (having an infinitely-large quantum capacitance). We
therefore conclude that the quantum capacitance-induced nonlinear
damping in the surface potential results in the observed salt
rejection.  Following the above discussions, we further investigate
the physical limits for the biased graphene nanopores. Specifically,
the salt rejection characteristics are controlled by: (i) the overlap
of EDL inside the nanopore, essentially controlled by two length
scales of $\lambda_{\mathrm{D}}$ and $r_{\mathrm{G}}$, and (ii) the
graphene surface potential $\psi_{\mathrm{G}}$ controlled by
$V_{\mathrm{G}}$ following Eq. \eqref{eq:Vg}. To this end, we
calculate $\psi_{\mathrm{G}}$ as a function of
$\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and $V_{\mathrm{G}}$ (\Fig{}
\ref{fig:5}a). Clearly, when the bulk concentration $c_{0}$ increases,
a higher $V_{\mathrm{G}}$ is required to reach the same level of
$\psi_{\mathrm{G}}$. Consequently, as revealed in \Fig{} \ref{fig:5}b,
the calculated salt rejection factor $\xi$ increases with both
$\lambda_{\mathrm{D}} / r_{\mathrm{G}}$ and $V_{\mathrm{G}}$, in line
with the experimentally observed $\xi - c_{0}$ dependence. We also
find the overall trends for $\psi_{\mathrm{G}}$ and $\xi$ with respect
to $\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ and $V_{\mathrm{G}}$ are very
similar. More importantly, based on our theoretical prediction, with
$V_{\mathrm{G}}$ = 1.25 V and $\lambda_{\mathrm{D}} / r_{\mathrm{G}}$
= 1.0, a high value of $\xi$ of $>$ 0.85 can be achieved.

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.95\linewidth]{img/fig5.pdf}
  \caption{\textbf{The Debye length effect on the degree of salt
      rejection}. \textbf{a}. Calculated contour plot of the surface
    potential of graphene $\psi_{\mathrm{G}}$ as a function of
    $\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ and $V_{\mathrm{G}}$. When
    the Debye length reduces, a higher $V_{\mathrm{G}}$ is required to
    maintain the same surface $\psi_{\mathrm{G}}$
    level. \textbf{b}. Calculated 2D contour plot of theoretically
    predicted $\xi$ as a function of
    $\lambda_{\mathrm{D}}/r_{\mathrm{G}}$ and
    $V_{\mathrm{G}}$. Clearly, a higher degree of salt rejection can
    be achieved by having a higher $V_{\mathrm{G}}$ and
    $\lambda_{\mathrm{D}}/r_{\mathrm{G}}$}
  \label{fig:5}
\end{figure}
